---
title: "Lab 4 Assignment"
author: "ChanceYan"
date: "2025-02-18"
output: html_document
---

Load data into R
```{r, include = TRUE}
canopy <- read.csv(file = "main_analysis_table_x-site_redlining_canopy.csv")
head(canopy)
```


Show what our data looks like:
```{r, include = TRUE}
table(canopy$city,canopy$holc_grade)
xtabs(~city+holc_grade, data=canopy) # another option
```


Combine Boston and NYC into one:
```{r, include = TRUE}
nyc <- c("Bronx", "Brooklyn", "Manhattan", "Queens", "Staten Island")
bos <- c("Boston", "Cambridge", "Chelsea")
canopy$city <- ifelse(canopy$city %in% nyc, "NYC",
ifelse(canopy$city %in% bos, "Boston", canopy$city))
remove(nyc, bos)
```



Plotting our data to see what it looks like inititally:
```{r, include = TRUE}
library(ggplot2)
## set colors for plotting
holc_col <- c('#92BC6B', '#92C7C9', '#E7DC6B', '#E47D67')
holc_lab <- c('A - "Best"',
'B - "Still Desirable"',
'C - "Definitely Declining"',
'D - "Hazardous"')

#Histogram
hist(canopy$Can_P, breaks=seq(0,100,5), cex.axis=0.8,
xlab="Canopy Cover (%)", main="")

#Boxplots
ggplot(canopy, aes(x = holc_grade, y = Can_P, fill = holc_grade)) +
geom_boxplot() +
scale_fill_manual('HOLC Grade', values = holc_col, labels = holc_lab) +
labs(title = "Canopy cover by HOLC grade",
y = "Canopy Cover (%)", x = "HOLC grade") +
theme_classic()

ggplot(canopy, aes(x = city, y = Can_P)) +
geom_boxplot() +
labs(title = "Canopy cover by city",
y = "Canopy Cover (%)", x = "City") +
theme_classic()
```

Removing cities with less than 150 observations
```{r, include = TRUE}
library(tidyverse)
## filter
canopy_s <- canopy %>%
group_by(city) %>%
filter(n() >= 150) %>%
ungroup()
## new table
table(canopy_s$city,canopy_s$holc_grade)
```

Now, let’s continue data exploration with the new simplified dataset. Try to create the
figures below by modifying your code from above.
```{r, include = TRUE}
#Histogram
hist(canopy_s$Can_P, breaks=seq(0,100,5), cex.axis=0.8,
xlab="Canopy Cover (%)", main="")

#Boxplots
ggplot(canopy_s, aes(x = holc_grade, y = Can_P, fill = holc_grade)) +
  geom_boxplot() +
  facet_wrap(~city) +
  ggtitle("Canopy Cover by HOLC grade (limited to cities with more than 150 observations)") +
  theme_classic() + 
  xlab("HOLC grade") + 
  ylab("Canopy Cover (%)") +
  guides(fill=guide_legend(title="HOLC grade"))
```



Looking at it by mean
```{r, include = TRUE}
## Mean canopy cover plot
canopy_mean <- canopy_s %>% group_by(city, holc_grade) %>%
summarise(Can_Pm = mean(Can_P))
ggplot(canopy_mean, aes(x = city, y = Can_Pm, group = holc_grade,
color = holc_grade)) +
geom_line() +
scale_color_manual('HOLC Grade', values = holc_col) +
labs(title = "Canopy cover by city",
y = "Canopy Cover (%)", x = "City") +
theme_classic()
```


Quiz yourself: (1pt) What do you notice about the raw data? How well do you think
our data will meet the assumptions of the linear model (we will again assume our data are
independent and measured without error)?

The raw data is 